Hierarchical Principal Manifold Estimation

### 2023-08-15 paper/main.tex
complete varying-coefficient model algorithm

The coefficients found at each time point $t$, $\mathbf{s'}_{t}$ and $\mathbf{\alpha'}_{t}$ are concatenated in matrix $\mathbf{\omega}_{t} = \left(\mathbf{s'}_{t}^T, \mathbf{\alpha'}_{t}^T\right)$. This coefficient matrix is then used as the output value in a cubic smoothing spline function using the time points as input values. This function takes the form
\begin{equation}
  g_{\gamma}(t) = \sum_{i=1}^{T}\delta_i\|t - t_i\|^{3} + \sum_{p=1}^{2}\tau_p \phi(t)_p. \label{eq:7}
\end{equation}
We define matrix $\mathbf{A}$ as $\mathbf{A}_{qr} = \|t_q - t_r\|^{3}$, matrix $\mathbf{T}$ as $\mathbf{T}_{pr} = \phi_p(t_r)$, $\mathbf{\delta} = \left(\delta_1, \dots, \delta_T\right)^{T}$, and $\mathbf{\tau} = \left(\tau_1, \tau_2\right)^{T}$. Then, the coefficients of $g_{\gamma}(t)$, given a value of $\gamma$, can be found by solving
\begin{equation}
  \left(
  \begin{array}{cc}
    \mathbf{A} + \gamma\mathbf{I} & \mathbf{T}^{T} \\
    \mathbf{T} & \mathbf{0}
  \end{array}
  \right)\left(
  \begin{array}{c}
    \mathbf{\delta} \\
    \mathbf{\tau}
  \end{array}
  \right) = \left(
  \begin{array}{c}
    \mathbf{\omega} \\
    \mathbf{0}
  \end{array}
  \right). \label{eq:8}
\end{equation}

As with the tuning process for $\lambda$, the optimal value of $\gamma$ can be identified using cross-validation, with either leave-one-out or $k$-folds cross-validation being appropriate depending on the number of time points available. With a proper estimate of the optimum $\gamma$, we may denote the coefficients estimated by $g(t)$ as $g(t) = \Omega(t) = \left(\mathbf{s}_{\gamma}(t)^{T}, \mathbf{\alpha}_{\gamma}(t)^{T}\right)^{T}$. Therefore, the varying-coefficient models have the following outputs:
\begin{equation}
  f(\mathbf{y}, t) = \sum_{k=1}^{m'}\mathbf{s}_{\gamma}(t)_l \eta_{4-d}\left(\|\mathbf{y} - \mathbf{y'}_k\|\right) + \sum_{p=1}^{d + 1}\mathbf{\alpha}_{\gamma}(t)_p \phi_p(\mathbf{y}). \label{eq:9}
\end{equation}
The fitting process for the varying-coefficient smoothing spline model is given formally as Algorithm \ref{alg:varying_spline}.

@@

\For{$t = 1, 2, \dots, T$} {
  Set $U_t$ equal to the number of images included in the data at time $t$\;
  \For{$l = 1, 2, \dots, L$} {
    Solve equation \ref{eq:5} to estimate $f_{\lambda_l, t}(\mathbf{y})$ with inputs $\mathbf{x}_t$, $\mathbf{y}_t$, and $w_t$\;
    \For{$u = 1, 2, \dots, U_t$} {
      Solve equation \ref{eq:5} to estimate $f^{(-u)}_{\lambda_l, t}(\mathbf{y})$ with inputs $\mathbf{x}^{(-u)}_t$, $\mathbf{y}^{(-u)}_t$, $w^{(-u)}_t$\;
      Estimate $MSE(\lambda_{l})_u$
    }
  }
}
\end{algorithm}

>>>>

The coefficients found at each time point $t$, $\mathbf{s'}_{t}$ and $\mathbf{\alpha'}_{t}$ are concatenated in matrix $\mathbf{\omega}_{t} = \left(\mathbf{s'}_{t}^T, \mathbf{\alpha'}_{t}^T\right)$. This coefficient matrix is then used as the output value in a cubic smoothing spline function using the time points as input values. This function takes the form
\begin{equation}
  g_{\gamma}(t) = \sum_{i=1}^{T}\delta_i\|t - t_i\|^{3} + \sum_{p=1}^{2}\tau_p \phi(t)_p. \label{eq:8}
\end{equation}
We define matrix $\mathbf{A}$ as $\mathbf{A}_{qr} = \|t_q - t_r\|^{3}$, matrix $\mathbf{T}$ as $\mathbf{T}_{pr} = \phi_p(t_r)$, $\mathbf{\delta} = \left(\delta_1, \dots, \delta_T\right)^{T}$, and $\mathbf{\tau} = \left(\tau_1, \tau_2\right)^{T}$. Then, the coefficients of $g_{\gamma}(t)$, given a value of $\gamma$, can be found by solving
\begin{equation}
  \left(
  \begin{array}{cc}
    \mathbf{A} + \gamma\mathbf{I} & \mathbf{T}^{T} \\
    \mathbf{T} & \mathbf{0}
  \end{array}
  \right)\left(
  \begin{array}{c}
    \mathbf{\delta} \\
    \mathbf{\tau}
  \end{array}
  \right) = \left(
  \begin{array}{c}
    \mathbf{\omega} \\
    \mathbf{0}
  \end{array}
  \right). \label{eq:9}
\end{equation}

As with the tuning process for $\lambda$, the optimal value of $\gamma$ can be identified using cross-validation, with either leave-one-out or $k$-folds cross-validation being appropriate depending on the number of time points available. With a proper estimate of the optimum $\gamma$, we may denote the coefficients estimated by $g(t)$ as $g(t) = \Omega(t) = \left(\mathbf{s}_{\gamma}(t)^{T}, \mathbf{\alpha}_{\gamma}(t)^{T}\right)^{T}$. Therefore, the varying-coefficient models have the following outputs:
\begin{equation}
  f(\mathbf{y}, t) = \sum_{k=1}^{m'}\mathbf{s}_{\gamma}(t)_l \eta_{4-d}\left(\|\mathbf{y} - \mathbf{y'}_k\|\right) + \sum_{p=1}^{d + 1}\mathbf{\alpha}_{\gamma}(t)_p \phi_p(\mathbf{y}). \label{eq:10}
\end{equation}
The fitting process for the varying-coefficient smoothing spline model is given formally as Algorithm \ref{alg:varying_spline}.

@@

\For{$t = 1, 2, \dots, T$} {
  Set $U_t$ equal to the number of images included in the data at time $t$\;
  \For{$l = 1, 2, \dots, L$} {
    Solve equation \ref{eq:5} to estimate $f_{\lambda_l, t}(\mathbf{y})$ with inputs $\mathbf{x}_t$, $\mathbf{y}_t$, and $w_t$\;
    \For{$u = 1, 2, \dots, U_t$} {
      Solve equation \ref{eq:5} to estimate $f^{(-u)}_{\lambda_l, t}(\mathbf{y})$ with inputs $\mathbf{x}^{(-u)}_t$, $\mathbf{y}^{(-u)}_t$, $w^{(-u)}_t$\;
      Estimate $MSE(\lambda_{l})_u$ using equation \ref{eq:6}\;
    }
    Set $MSE(\lambda_l) = \frac{1}{U}\sum_{u=1}^{U}MSE(\lambda_l)_u$\;
  }
  $\lambda_t = \arg \min_{\lambda} MSE(\lambda)$\;
  $f_t(\mathbf{y}) = f_{\lambda_t}(\mathbf{y})$\;
}

  Let $m' = \max(m_{ij})$, $\left\{\mathbf{y}'_{k}\right\}_{k=1}^{m'}$ be a grid spanning the range of values of $\left\{\mathbf{y}\right\}$\;
  \For{$t = 1, 2, \dots, T$} {
    Compute $\left\{\mathbf{x'}_{k, t}\right\}_{k=1}^{m'} = \left\{f_t(\mathbf{y'}_k)\right\}$\;
    Solve equation \ref{eq:7} to compute $\mathbf{s'}_t$ and $\mathbf{\alpha'}_t$\;
    Set $\mathbf{\omega}_t = \left[\mathbf{s'}_t^{T}, \mathbf{\alpha'}_t^{T}\right]$\;
  }
  \For{$p = 1, 2, \dots, P$} {
    Compute $g_{\gamma_l}(t)$ by solving equation \ref{eq:9}\;
    \For{$t = 1, 2, \dots, T$} {
      Compute $g_{\gamma_l}^{(t)}$ by solving \ref{eq:9}\;
    }
    Estimate $MSE(\gamma_l)$ using \ref{eq:10}\;
  }
  $\hat{\gamma} = \arg \min_{\gamma} MSE(\gamma)$\;
  $\hat{f}(\mathbf{y}, t) = f_{\hat{\gamma}}(\mathbf{y}, t)$, where the form of $\hat{f}(\mathbf{y}, t)$ is given in \ref{eq:11}
\end{algorithm}
